---
title: "README for SMI205 Assessmnet 2"
author: "200382029"
date: "2023-05-22"
output:
  github_document:
    toc: true
    toc_depth: 2
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# 1. Replication report

The `SMI205_Assessment2_Template.Rmd` is enlisted within the repository showing all steps taken including cod, graphs and descriptions. It beings with workspace setup information such as libraries and environment. The template then includes academic research regarding similar issues and further followed with information regarding where and how the data was installed and changes to variables. Outputs for each of the three regressions are shown and then followed with regression models and other outputs showing distrust in government and a conclusion summarizing the findings.
  
The work has then been published  as a reproducible report on [Rpubs.com](https://rpubs.com/a200382029/1047951) website and an online repository project was formed on Github.


## 1.1. Your Workflow 

This workflow includes multiple folders: two readme files, the work project, template and pre-registration form

## 1.2. Re-publishing data and code

Data has not been republished or created


# 2. Rpubs version of the template

HTML version of the template is published here:
https://rpubs.com/a200382029/1047951

# 3. `renv` package

The Template.Rproj was created without using `renv` package, yet, you could consider using it. The `renv` package helps you to create a reproducible environment for your R project. Read more here: https://rstudio.github.io/renv/.

It saves information about R and loaded packages. So if you later (after any R updates or changes in the packages) or other people open your Rproj, it will install the same libraries, and will not use the central libraries installed on a computer ([Joseph 2022](https://medium.com/@adrian.joseph/renv-make-r-environment-reproducible-414d88c683aa )). 

## 3.1. How to work with `renv` package



+ Call `renv::init()` to initialise a new project-local environment with a private R library.
+ Call `renv::snapshot()` to save the state of the project library to the lockfile (called renv.lock).
+ Call `renv::restore()` to recall the packages and version the last time you called `renv::snapshot()`.

They are called automatically when you work in RProj and initiate `renv` at the start.

## 3.2. Trouble shooting with `renv` package

This is a useful guide by Joseph A. (2022). renv: make R environment reproducible. URL: https://medium.com/@adrian.joseph/renv-make-r-environment-reproducible-414d88c683aa (accessed 09/02/2023).

# 4. Transparency checklist

+ All relevant files have been organised and saved in folders
+ Are all files, variables and R objects have been named and organised (in the template folder)
+ A READme file has been included
+ This project has been tested to run on a different device
+ Information about your workspace specifications have been saved in your READme file
+ No sensitive information has been included in the Github  
 
# 5. My enviroment

Below you can find information about versions of R and specific packages used to create this project. 

```{r envir}
sessionInfo()
```

